% ---- Abstract ---------------------------------------------------------------
% Target length: 150–250 words. Write last, once results exist.
% Key elements: problem, gap, approach, dataset, results, significance.

Scalable Vector Graphics (SVG) is the native format for resolution-independent
icons, yet generating high-quality monochrome SVG icons from natural language
descriptions remains an unsolved problem. Professional icon design follows a
strict convention: compact, single-colour shapes using the \texttt{currentColor}
CSS property, enabling immediate brand recolouring across any design system
without modifying the source file.

Existing text-to-SVG approaches target multi-colour illustrations, logos, and
raster-to-vector conversions. None are optimised for the compact, semantically
precise, single-colour structure required by production icon workflows --- leaving
a clear gap for designers and developers who need a lightweight, pixel-perfect
icon generator that works across organisations and brands.

We present \textbf{\method}, a text-to-SVG model focused exclusively on
monochrome icon generation. Our approach frames icon generation as a structured
code generation task and fine-tunes a pre-trained code language model
(Qwen3-Coder-30B) using Low-Rank Adaptation (LoRA) on a purpose-built dataset of
natural language and SVG icon pairs.

To support this, we introduce the \textbf{\dataset{}} corpus: \num{275912}
SVG icons sourced from the Iconify open-source icon ecosystem, filtered by
license, normalised to a consistent monochrome format with \texttt{currentColor},
and annotated with natural language descriptions generated by a vision language
model (Qwen2.5-VL-7B).

\todo{Fill in: key quantitative results once experiments are complete.}

\method{} generates pixel-perfect, immediately renderable, instantly recolourable
SVG icons in under one second. Code, model weights, and the \dataset{} corpus
are released publicly under open-source licenses.
